---
title: "NBA Shot Analysis"
subtitle: "By Vedhant Sarma"
output:
  pdf_document:
    extra_dependencies: ["bm", "bbm", "amsmath", "physics"]
urlcolor: blue
---
```{r setup, include = FALSE}
library(knitr)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  tidy = "styler"
)
```

$\;$
$\;$
$\;$
$\;$

The data in `NBA_2024_Shots.csv` includes $27$ variates about all $218,701$ shots taken in the 2023-24 National Basketball Association regular season. This data was scraped from [nba.com](https://www.nba.com/) and is available on [Kaggle](https://www.kaggle.com/datasets/mexwell/nba-shots/). For every shot, the following variates will be of interest.

| Variable               | Description                                                 |
|------------------------|-------------------------------------------------------------|
| `TEAM_NAME`            | The name of the team of the player who took the shot.       |
| `TEAM_NAME_ABBREV`     | A 3-letter abbreviation for the name of the team of the player who took the shot.       |
| `PLAYER_NAME`          | The name of the player who took the shot.                   |
| `HOME_TEAM`            | The home team in the game in which the shot was taken. This is recorded as the team's 3-letter abbreviation.      |
| `AWAY_TEAM`            | The away team in the game in which the shot was taken. This is recorded as the team's 3-letter abbreviation.       |
| `SHOT_MADE`            | A boolean variate indicating whether the shot was successful (`TRUE`) or unsuccessful (`FALSE`). |
| `LOC_X`                | The $x$ coordinate of the shot in the $(x,y)$ plane of the court. This dimension corresponds to the width of the court, which is recorded in feet and coded between `(-25,25)`.|
| `LOC_Y`                | The $y$ coordinate of the shot in the $(x,y)$ plane of the court. This dimension corresponds to the length of the court, which is recorded in feet and coded between `(0,94)`.|


Below is a visualization of the court, which should help to interpret `LOC_X` and `LOC_Y`. The target hoop is the one located at (0,4).

```{r, echo=FALSE, fig.height=5.5, fig.align='center'}
draw_court <- function(add = FALSE, lwd = 1, col = "black", axes = FALSE) {
  if(!add) {
    plot(0,0, col = "white", asp = 1,
         xlim = c(-30,30), xaxt = "n", xlab = "",
         ylim = c(0,95), yaxt = "n", ylab = "")
    if(axes) {
      axis(side = 1, cex.axis = 0.75)
      axis(side = 2, cex.axis = 0.75)
    }
  }
  # baselines
  segments(x0 = -25, y0 = 0, x1 = 25, y1 = 0, lwd = lwd, col = col)
  segments(x0 = -25, y0 = 94, x1 = 25, y1 = 94, lwd = lwd, col = col)
  # sidelines
  segments(x0 = -25, y0 = 0, x1 = -25, y1 = 94, lwd = lwd, col = col)
  segments(x0 = 25, y0 = 0, x1 = 25, y1 = 94, lwd = lwd, col = col)
  # half court
  segments(x0 = -25, y0 = 47, x1 = 25, y1 = 47, lwd = lwd, col = col)
  lines(x = seq(-6,6,0.01), y = sqrt(36-seq(-6,6,0.01)^2)+47, lwd = lwd, col = col)
  lines(x = seq(-6,6,0.01), y = -sqrt(36-seq(-6,6,0.01)^2)+47, lwd = lwd, col = col)
  # 3-point arcs
  lines(x = seq(-22,22,0.01), y = sqrt(23.9^2-seq(-22,22,0.01)^2)+4.7, lwd = lwd, col = col)
  segments(x0 = -22, y0 = 0, x1 = -22, y1 = 14, lwd = lwd, col = col)
  segments(x0 = 22, y0 = 0, x1 = 22, y1 = 14,lwd = lwd, col = col)
  lines(x = seq(-22,22,0.01), y = -sqrt(23.9^2-seq(-22,22,0.01)^2)+89.3, lwd = lwd, col = col)
  segments(x0 = -22, y0 = 80, x1 = -22, y1 = 94, lwd = lwd, col = col)
  segments(x0 = 22, y0 = 80, x1 = 22, y1 = 94, lwd = lwd, col = col)
  # keys
  lines(x = seq(-6,6,0.01), y = sqrt(36-seq(-6,6,0.01)^2)+15, lwd = lwd, col = col)
  lines(x = seq(-6,6,0.01), y = -sqrt(36-seq(-6,6,0.01)^2)+15, lwd = lwd, lty = 3, col = col)
  segments(x0 = -8, y0 = 0, x1 = -8, y1 = 15, lwd = lwd, col = col)
  segments(x0 = 8, y0 = 0, x1 = 8, y1 = 15, lwd = lwd, col = col)
  segments(x0 = -8, y0 = 15, x1 = 8, y1 = 15, lwd = lwd, col = col)
  lines(x = seq(-6,6,0.01), y = -sqrt(36-seq(-6,6,0.01)^2)+79, lwd = lwd, col = col)
  lines(x = seq(-6,6,0.01), y = sqrt(36-seq(-6,6,0.01)^2)+79, lwd = lwd, lty = 3, col = col)
  segments(x0 = -8, y0 = 94, x1 = -8, y1 = 79, lwd = lwd, col = col)
  segments(x0 = 8, y0 = 94, x1 = 8, y1 = 79, lwd = lwd, col = col)
  segments(x0 = -8, y0 = 79, x1 = 8, y1 = 79, lwd = lwd, col = col)
  # baskets
  points(x = 0, y = 4, pch = 16, col = col)
  segments(x0 = -2, y0 = 3, x1 = 2, y1 = 3, lwd = lwd, col = col)
  points(x = 0, y = 90, pch = 16, col = col)
  segments(x0 = -2, y0 = 91, x1 = 2, y1 = 91, lwd = lwd, col = col)
}
draw_court(axes = TRUE)
```

\newpage

*Field goal percentage* ($FG\%$) is a basketball statistic defined as $$FG\%=\frac{\#\text{ shots made}}{\#\text{ shots taken}}\times 100$$ which measures how well a player (or team) shoots the ball. Values of $FG\%$ closer to 100 indicate a better player (or team) and $FG\%$ values closer to 0 indicate a worse player (or team). Given a population $\mathcal{P}$ of shots, this attribute can be more precisely stated as $$FG\% = a(\mathcal{P})=\frac{100}{N}\sum_{u\in\mathcal{P}}\mathbb{I}\{\text{shot}~u~\text{was made}\}.$$ For the purpose of this question we will assume the shots taken in the 2023-2024 regular season constitute the population of interest $\mathcal{P}$. We then use the Horvitz-Thompson estimation framework to estimate the league-wide $FG\%$ in the 2023-2024 NBA regular season based on a sample of these shots. You will also study the sampling distribution of this estimator, and the coverage of associated confidence intervals.

Here, we take a *simple random sample without replacement* $\mathcal{S}$ of size $n=1000$ from the population.
      ```{r, eval=FALSE}
      nba <- read.csv("NBA_2024_Shots.csv", header = TRUE)
      set.seed(341)
      srswor_indices <- sample(x = nrow(nba), size = 1000)
      srswor_sample <- nba[srswor_indices,]
      ```
```{r}
nba <- read.csv("NBA_2024_Shots.csv", header = TRUE)
set.seed(341)
srswor_indices <- sample(x = nrow(nba), size = 1000)
srswor_sample <- nba[srswor_indices,]
```
Using this sample, we can calculate the Horvitz-Thompson estimate of the league-wide field goal percentage $FG\%$.
      
```{r}
shots_made <- srswor_sample$SHOT_MADE

N <- nrow(nba)  # Total population size
n <- 1000

pi_u <- n / N   # Inclusion probability

ht_fg_percentage <- 100 * sum(shots_made / pi_u) / N
cat("FG % =", ht_fg_percentage, "%")
```

Using this sample, calculate the following three 95% confidence intervals for the league-wide field goal percentage $FG\%$:
            
```{r}
est_var_ht <- function(t_u, pi_u, pi_uv) {
delta <- pi_uv - outer(pi_u, pi_u)
var_hat <- sum((delta/pi_uv) * outer(t_u/pi_u, t_u/pi_u))
return(var_hat)
}

pi_u <- rep(n/N, n)
#Joint Inclusion Matrix
pi_uv <- matrix(((n * (n-1)) / (N * (N-1))), nrow = n, ncol = n)
diag(pi_uv) <- pi_u

# Calculate both confidence intervals, upper and lower
CI_Int <- ht_fg_percentage + 1.96 * c(-1, 1) * sqrt(est_var_ht((100*shots_made)/N, pi_u, pi_uv))

cat("HT Estimate of FG%: [", CI_Int, "]")
```
            
```{r}
B <- 1000
set.seed(341)
bootstrap_estimates <- numeric(B)
pi_u <- matrix(ht_fg_percentage, nrow = n, ncol = n)

#Bootstrap Sampling
for (i in 1:B) {
  resample_indices <- sample(srswor_indices, size = 1000, replace = TRUE)
  resample <- nba[resample_indices,]
  y_resample <- resample$SHOT_MADE
  bootstrap_estimates[i] <- sum(y_resample / pi_u) / N
}

# Bootstrap variance and confidence intervals
boot_variance <- var(bootstrap_estimates)
ci_normal <- ht_fg_percentage + 1.96 * c(-1, 1) * sqrt(boot_variance)
ci_percentile <- quantile(bootstrap_estimates, probs = c(0.025, 0.975))

cat("95% Normal Confidence Interval: [", ci_normal[1], ",", ci_normal[2], "]\n")
cat("95% Percentile Confidence Interval: [", ci_percentile[1], ",", ci_percentile[2], "]")
```

\newpage

Exploring the performance of several confidence intervals for $FG\%$ as a function of sample size. Consider the sample sizes $n \in \{50, 100, 500, 1000\}$. 

For each sample size $n$, we take 500 SRSWOR samples from the population. 
        * For each of the 500 samples of a given size, we calculate:
            - Horvitz-Thompson estimate of $FG\%$ 
            - Three 95% confidence intervals for $FG\%$ from above
        * For each sample size $n$, we use the 500 HT estimates to estimate the sampling bias, sampling variance, sampling MSE of the estimator.
        * For each sample size $n$, we also use the 500 confidence intervals of each type to estimate their coverage.
        * Constructing the following six line-plots (laid out in a $2 \times 3$ grid). 
            * Plot 1: Bias vs. $n$, with y-axis limits [-5,5] and a dashed horizontal line at 0. 
            * Plot 2: Variance vs. $n$, with y-axis limits [0,50].
            * Plot 3: MSE vs. $n$, with y-axis limits [0,50].
            * Plot 4: Coverage for the naive Horvitz-Thompson CI vs $n$, with y-axis limits $[0,1]$ and a dashed horizontal line at 0.95. 
            * Plot 5: Coverage for the naive normal theory bootstrap CI vs $n$, with y-axis limits $[0,1]$ and a dashed horizontal line at 0.95. 
            * Plot 6: Coverage for the percentile method bootstrap CI vs $n$, with y-axis limits $[0,1]$ and a dashed horizontal line at 0.95. 
    
```{r}
nba <- read.csv("NBA_2024_Shots.csv")
shots_made <- nba$SHOT_MADE
 # True HT FG%
true_fg_percentage <- sum(shots_made) / nrow(nba) * 100

sample_sizes <- c(50, 100, 500, 1000)
B <- 50  # Number of bootstrap samples
M <- 50  # Number of SRSWOR samples
N <- nrow(nba)

results <- list()

set.seed(341)
for (n in sample_sizes) {
  HT_estimates <- numeric(M)
  naive_coverage <- numeric(M)
  normal_bootstrap_coverage <- numeric(M)
  percentile_bootstrap_coverage <- numeric(M)
  
  for (i in 1:M) {
    # Take SRSWOR sample
    sample_indices <- sample(N, size = n, replace = FALSE)
    sample_data <- nba[sample_indices,]
    
    # HT estimate
    pi_u <- rep(n / N, n)
    HT_estimates[i] <- sum(sample_data$SHOT_MADE / pi_u) / N * 100
    
    # Bootstrap resampling
    bootstrap_estimates <- numeric(B)
    for (b in 1:B) {
      resample_indices <- sample(sample_indices, size = n, replace = TRUE)
      resample_data <- nba[resample_indices,]
      bootstrap_estimates[b] <- sum(resample_data$SHOT_MADE / pi_u) / N * 100
    }
    
    # Confidence Intervals
    naive_CI <- HT_estimates[i] + c(-1.96, 1.96) * sd(HT_estimates) / sqrt(n)
    normal_bootstrap_CI <- mean(bootstrap_estimates) + c(-1.96, 1.96) * sd(bootstrap_estimates)
    percentile_CI <- quantile(bootstrap_estimates, probs = c(0.025, 0.975))
    
    # Check Coverage
    naive_coverage[i] <- (naive_CI[1] <= true_fg_percentage) && (naive_CI[2] >= true_fg_percentage)
    normal_bootstrap_coverage[i] <- (normal_bootstrap_CI[1] <= true_fg_percentage) && (normal_bootstrap_CI[2] >= true_fg_percentage)
    percentile_bootstrap_coverage[i] <- (percentile_CI[1] <= true_fg_percentage) && (percentile_CI[2] >= true_fg_percentage)
  }
  
  # Calculate Bias, Variance, MSE, and Coverage
  bias <- mean(HT_estimates) - true_fg_percentage
  variance <- var(HT_estimates)
  mse <- bias^2 + variance
  
  results[[as.character(n)]] <- list(
    bias = bias,
    variance = variance,
    mse = mse,
    naive_coverage = mean(naive_coverage),
    normal_bootstrap_coverage = mean(normal_bootstrap_coverage),
    percentile_bootstrap_coverage = mean(percentile_bootstrap_coverage), 
    HT_estimates = mean(HT_estimates),
    HT_estimate_CI = naive_CI,
    Bootstrap_CI = normal_bootstrap_CI,
    Percentile_CI = percentile_CI)
}
print(results)
```

```{r}
par(mfrow = c(2, 3))

bias_values <- c(0, 0, 0, 0)
variance_values <- c(0, 0, 0, 0)
MSE_values <- c(0, 0, 0, 0)
N_Cvg_values <- c(0, 0, 0, 0)
Norm_Boot_values <- c(0, 0, 0, 0)
P_Boot_values <- c(0, 0, 0, 0)

#Iterate over values for Plotting
for (i in 1:4) {
  bias_results <- results[[i]][[1]]
  bias_values[i] <- bias_results
  
  variance_results <- results[[i]][[2]]
  variance_values[i] <- variance_results
  
  MSE_results <- results[[i]][[3]]
  MSE_values[i] <- MSE_results
  
  N_Cvg_results <- results[[i]][[4]]
  N_Cvg_values[i] <- N_Cvg_results
  
  Norm_Boot_results <- results[[i]][[5]]
  Norm_Boot_values[i] <- Norm_Boot_results
  
  P_Boot_results <- results[[i]][[6]]
  P_Boot_values[i] <- P_Boot_results
}

plot(x = sample_sizes, y = bias_values, type = "l", ylim = c(-5, 5), main = "Bias vs. n", ylab = "Bias", xlab = "n")
abline(h = 0, lty = 2)

plot(x = sample_sizes, y = variance_values , type = "b", ylim = c(0, 50), main = "Variance vs. n", ylab = "Variance", xlab = "n")

plot(x = sample_sizes, y = MSE_values , type = "b", ylim = c(0, 50), main = "MSE vs. n", ylab = "MSE", xlab = "n")

plot(x = sample_sizes, y = N_Cvg_values , type = "b", ylim = c(0, 1), main = "Naive CI Coverage vs. n", ylab = "Coverage", xlab = "n")
abline(h = 0.95, lty = 2)

plot(x = sample_sizes, y = Norm_Boot_values, type = "b", ylim = c(0, 1), main = "Normal Boot CI Coverage vs. n", ylab = "Coverage", xlab = "n")
abline(h = 0.95, lty = 2)

plot(x = sample_sizes, y = P_Boot_values, type = "b", ylim = c(0, 1), main = "Percentile Boot CI Coverage vs. n", ylab = "Coverage", xlab = "n")
abline(h = 0.95, lty = 2)

```

\newpage

In the NBA, when a team plays on their home court, this is considered a "home" game, and when they play on their opponent's home court, this is considered an "away" game. It is often thought teams perform better while playing at home versus away, and this is referred to as the so-called "home court advantage". Here, we will investigate whether the field goal percentage $FG\%$ differs dramatically for home teams versus away teams.

Let us define $\mathcal{P}_1$ as the population of all NBA shots ever taken by home teams and $\mathcal{P}_2$ as the population of all NBA shots ever taken by away teams. The shots recorded in `NBA_2024_Shots.csv` are clearly a subset of $\mathcal{P}_1$ and $\mathcal{P}_2$ since they correspond only to shots taken in the 2023-2024 regular season. As such, the home team shots in `NBA_2024_Shots.csv` could be regarded as a sample $\mathcal{S}_1$ of $\mathcal{P}_1$, and the away team shots could be considered a sample $\mathcal{S}_2$ of $\mathcal{P}_2$. We use the randomization test to compare $\mathcal{P}_1$ and $\mathcal{P}_2$ by way of $\mathcal{S}_1$ and $\mathcal{S}_2$.

Calculating $FG\%_{home}$ and $FG\%_{away}$, the field goal percentages respectively calculated for 2023-2024 shots made by home teams ($\mathcal{S}_1$) versus 2023-2024 shots made by away teams ($\mathcal{S}_2$). We visualize these values and hence the `SHOT_MADE` distributions for home teams versus away teams by creating stacked bar plots shown below.

```{r}
nba <- read.csv("NBA_2024_Shots.csv")

team_mapping <- c(
  "ATL" = "Atlanta Hawks", "BOS" = "Boston Celtics", "BKN" = "Brooklyn Nets",
  "CHA" = "Charlotte Hornets", "CHI" = "Chicago Bulls", "CLE" = "Cleveland Cavaliers",
  "DAL" = "Dallas Mavericks", "DEN" = "Denver Nuggets", "DET" = "Detroit Pistons",
  "GSW" = "Golden State Warriors", "HOU" = "Houston Rockets", "IND" = "Indiana Pacers",
  "LAC" = "LA Clippers", "LAL" = "Los Angeles Lakers", "MEM" = "Memphis Grizzlies",
  "MIA" = "Miami Heat", "MIL" = "Milwaukee Bucks", "MIN" = "Minnesota Timberwolves",
  "NOP" = "New Orleans Pelicans", "NYK" = "New York Knicks", "OKC" = "Oklahoma City Thunder",
  "ORL" = "Orlando Magic", "PHI" = "Philadelphia 76ers", "PHX" = "Phoenix Suns",
  "POR" = "Portland Trail Blazers", "SAC" = "Sacramento Kings", "SAS" = "San Antonio Spurs",
  "TOR" = "Toronto Raptors", "UTA" = "Utah Jazz", "WAS" = "Washington Wizards"
)

# Convert abbreviations to full names
nba$HOME_TEAM_FULL <- team_mapping[nba$HOME_TEAM]
nba$AWAY_TEAM_FULL <- team_mapping[nba$AWAY_TEAM]

# Home and Away Shots
home_shots <- subset(nba, TEAM_NAME == HOME_TEAM_FULL)
away_shots <- subset(nba, TEAM_NAME == AWAY_TEAM_FULL)

home_fg_percentage <- sum(home_shots$SHOT_MADE) / nrow(home_shots) * 100
away_fg_percentage <- sum(away_shots$SHOT_MADE) / nrow(away_shots) * 100

cat("Home FG %:", home_fg_percentage, "%\n")
cat("Away FG %:", away_fg_percentage, "%")

home_shot_distribution <- table(factor(home_shots$SHOT_MADE, levels = c(0, 1)))
away_shot_distribution <- table(factor(away_shots$SHOT_MADE, levels = c(0, 1)))

# Combine data
shot_distribution <- rbind(home_shot_distribution, away_shot_distribution)
rownames(shot_distribution) <- c("Made", "Missed")
colnames(shot_distribution) <- c("Missed", "Made")

# Plot the stacked bar plot
barplot(
  shot_distribution, beside = FALSE, col = c("forestgreen", "firebrick"),
  main = "Home vs Away Shot Outcomes",
  xlab = "Location", ylab = "Number of Shots",
  names.arg = c("Home", "Away")
)

```
There seems to be a neglibible difference of field goal percentage indicating a small home court advantage, but not statistically significant

\begin{center}

\end{center}

Perfomring a hypothesis test using the discrepancy measure $$D(\mathcal{S}_1,\mathcal{S}_2) = \left|FG\%_{home} - FG\%_{away}\right|.$$
```{r}
set.seed(341)

diff_FG <- abs(home_fg_percentage - away_fg_percentage)
print(diff_FG)
```

Here we randomly mix $\mathcal{S}_1$ and $\mathcal{S}_2$ $M=1000$ times and construct a histogram of the 1000 $D(\mathcal{S}_1^\star,\mathcal{S}_2^\star)$ values. The vertical line represents the observed discrepancy calculated previously
    
```{r}
mix_randomly <- function(pop1, pop2) {
  n_pop1 <- nrow(pop1)
  n_pop2 <- nrow(pop2)
  full_pop <- rbind(pop1, pop2)
  rid_pop1 <- sample(1:(n_pop1 + n_pop2), n_pop1)
  new_pop1 <- full_pop[rid_pop1, ]
  new_pop2 <- full_pop[-rid_pop1, ]
  list(pop1 = new_pop1, pop2 = new_pop2)
}

# Perform 1000 random mixes
M <- 1000
discrepancies <- numeric(M)

set.seed(123)
for (i in 1:M) {
  mixed <- mix_randomly(home_shots, away_shots)
  FG_home_mixed <- sum(mixed$pop1$SHOT_MADE) / nrow(mixed$pop1) * 100
  FG_away_mixed <- sum(mixed$pop2$SHOT_MADE) / nrow(mixed$pop2) * 100
  discrepancies[i] <- abs(FG_home_mixed - FG_away_mixed)
}

# Plot the histogram
hist(discrepancies, breaks = 30, col = "blue", main = "Histogram of Discrepancies",
     xlab = "Discrepancies")
abline(v = diff_FG, col = "red", lwd = 2)
```

P-value associated with this test is:
```{r}
set.seed(341)
p_value <- mean(abs(discrepancies) >= diff_FG)
cat("P-value:", p_value, "\n")
```
    
The p-value being 0.001 indicates that there is strong evidence against $$H_0$$. Since there is strong evidence against $$S_1$$ and $$S_2$$ being drawn from the same population, it indicates there is some sort of home court as there is a difference between the two populations/samples of shots (Home vs Away). It coincides with what I drew from earlier where there is practically some sort of home court advantage from looking at the FG percentages

\newpage

Interest lies in estimating the probability of successfully making a shot, as a function of where on the court the shot is taken. Denote $$\theta(\mathbf{x}) = \text{Pr}(\text{shot is successful when taken from position }\mathbf{x})$$ as this probability, where $\mathbf{x}=(x,y)$ is some position on the court (i.e., $x\in[-25,25]$, $y\in[0,94]$). We will use 2-dimensional smoothing to estimate this probability over the entire court. The 218,701 shots from `NBA_2024_Shots.csv` will be used as a sample $\mathcal{S}$ of all shots ever taken in the NBA, and we will use this sample to perform the relevant estimation.

Scatter plot of $y=$ `LOC_Y` vs. $x=$ `LOC_X` using all $n=218,701$ shots in the sample $\mathcal{S}$ (i.e., all shots in `NBA_2024_Shots.csv`). Coloured points `adjustcolor(forestgreen, 0.2)` if the corresponding shot was successful and `adjustcolor(firebrick, 0.2)` if it was unsuccessful.

```{r}
#Colour in shot if missed or made respectively
nba_shots <- ifelse(nba$SHOT_MADE, 
                               adjustcolor("forestgreen", 0.2), 
                               adjustcolor("firebrick", 0.2))

draw_court <- function(add = FALSE, lwd = 1, col = "black", axes = FALSE) {
if(!add) {
plot(
x = 0, y = 0,
col = "white", asp = 1,
xlim = c(-25,25), xaxt = "n", xlab = "",
ylim = c(0,94), yaxt = "n", ylab = ""
)
if(axes) {
axis(side = 1, cex.axis = 0.75)
axis(side = 2, cex.axis = 0.75)
}
}
segments(x0 = -25, y0 = 0, x1 = 25, y1 = 0, lwd = lwd, col = col)
segments(x0 = -25, y0 = 94, x1 = 25, y1 = 94, lwd = lwd, col = col)
segments(x0 = -25, y0 = 0, x1 = -25, y1 = 94, lwd = lwd, col = col)
segments(x0 = 25, y0 = 0, x1 = 25, y1 = 94, lwd = lwd, col = col)
segments(x0 = -25, y0 = 47, x1 = 25, y1 = 47, lwd = lwd, col = col)
lines(x = seq(-6,6,0.01), y = sqrt(36-seq(-6,6,0.01)^2)+47, lwd = lwd, col = col)
lines(x = seq(-6,6,0.01), y = -sqrt(36-seq(-6,6,0.01)^2)+47, lwd = lwd, col = col)
lines(x = seq(-22,22,0.01), y = sqrt(23.9^2-seq(-22,22,0.01)^2)+4.7, lwd = lwd, col = col)
segments(x0 = -22, y0 = 0, x1 = -22, y1 = 14, lwd = lwd, col = col)
segments(x0 = 22, y0 = 0, x1 = 22, y1 = 14,lwd = lwd, col = col)
lines(x = seq(-22,22,0.01), y = -sqrt(23.9^2-seq(-22,22,0.01)^2)+89.3, lwd = lwd, col = col)
segments(x0 = -22, y0 = 80, x1 = -22, y1 = 94, lwd = lwd, col = col)
segments(x0 = 22, y0 = 80, x1 = 22, y1 = 94, lwd = lwd, col = col)
lines(x = seq(-6,6,0.01), y = sqrt(36-seq(-6,6,0.01)^2)+15, lwd = lwd, col = col)
lines(x = seq(-6,6,0.01), y = -sqrt(36-seq(-6,6,0.01)^2)+15, lwd = lwd, lty = 3, col = col)
segments(x0 = -8, y0 = 0, x1 = -8, y1 = 15, lwd = lwd, col = col)
segments(x0 = 8, y0 = 0, x1 = 8, y1 = 15, lwd = lwd, col = col)
segments(x0 = -8, y0 = 15, x1 = 8, y1 = 15, lwd = lwd, col = col)
lines(x = seq(-6,6,0.01), y = -sqrt(36-seq(-6,6,0.01)^2)+79, lwd = lwd, col = col)
lines(x = seq(-6,6,0.01), y = sqrt(36-seq(-6,6,0.01)^2)+79, lwd = lwd, lty = 3, col = col)
segments(x0 = -8, y0 = 94, x1 = -8, y1 = 79, lwd = lwd, col = col)
segments(x0 = 8, y0 = 94, x1 = 8, y1 = 79, lwd = lwd, col = col)
segments(x0 = -8, y0 = 79, x1 = 8, y1 = 79, lwd = lwd, col = col)
points(x = 0, y = 4, pch = 16, col = col)
segments(x0 = -2, y0 = 3, x1 = 2, y1 = 3, lwd = lwd, col = col)
points(x = 0, y = 90, pch = 16, col = col)
segments(x0 = -2, y0 = 91, x1 = 2, y1 = 91, lwd = lwd, col = col)
}
#Plot shots
plot(nba$LOC_X, nba$LOC_Y, col = nba_shots)
#Overlay Court
draw_court(add = TRUE)
```

Now let's create a heatmap and countours to visualize our estimates of the likelihood of making a shot from anywhere on the court

Steps:
    * Discretize the court into a $50\times 50$ grid (where each of the 2500 cells has the same size).
    * Denote the center of a grid cell as $\mathbf{x}_0$. Identify all shots taken within 10 feet in any direction of $\mathbf{x}_0$. Denote this neighbourhood of shots by $\mathcal{N}_0$.
    * Calculate $$\hat{\theta}(\mathbf{x}_0)=\frac{\sum_{u\in\mathcal{N}_0}w_u\mathbb{I}\{\text{shot}~u~\text{was made}\}}{\sum_{u\in\mathcal{N}_0}w_u},$$ the weighted average of `SHOT_MADE` for all shots $u\in\mathcal{N}_0$, where the weight for shot $u$ is defined as $$w_u = \frac{1}{1+\lVert \mathbf{x}_0 - \mathbf{x}_u\rVert}.$$ This weighted average estimates $\theta(\mathbf{x})$, the probability that a shot is successful, when taken from position $\mathbf{x}=\mathbf{x}_0$.
    
```{r}
cells <- 50
radius <- 10
width <- seq(-24, 25, legnth.out = cells)
height <- seq(0, 94, length.out = cells)
col_fun <- colorRampPalette(c("firebrick", "forestgreen"))
theta_hat <- matrix(0, nrow = cells-1, ncol = cells-1)
draw_court(add = FALSE)

for(i in 1:(cells-1)) {
  for (j in 1:(cells-1)) {
    x_0 <- (width[i] + width[i+1]) / 2
    y_0 <- (height[j] + height[j+1]) / 2 
    dist <- sqrt((x_0 - nba$LOC_X)^2 + (y_0 - nba$LOC_Y)^2)
    N_0 <- which(dist <= radius)
    if (length(N_0) > 0) {
    theta_est <- weighted.mean(nba[N_0, "SHOT_MADE"], w = 1 / (1 + dist[N_0]))
    }
    rect(xleft = width[i], xright = width[i+1], ybottom = height[j], ytop = height[j+1], 
         col = col_fun(10000)[1 + floor(10000 * theta_est)], 
         border = col_fun(10000)[1 + floor(10000 * theta_est)]
    )
    theta_hat[i, j] <- theta_est
  }
}
contour(x = width[-cells] + diff(width)[i] / 2, 
        y = height[-cells] + diff(height)[i] / 2, 
        z = theta_hat, add = TRUE, col = "grey", levels = seq(0, 0.5, 0.05))
draw_court(add = TRUE, lwd = 2)
```

Now, to look at the most common spots on the court and calculate $\hat{\theta}(\mathbf{x}_0)$ at each of the specific coordinates
    * Free throw position: $\mathbf{x}_0=(0,16)$
    * Top of the 3-point arc: $\mathbf{x}_0=(0,30)$
    * Center court: $\mathbf{x}_0=(0,47)$

Let's also calculate a 95% confidence interval for $\theta(\mathbf{x}_0)$ using the bootstrap percentile method
    
```{r}
# Initialize positions of interest
positions <- list(
  free_throw = c(0, 16),
  three_point_arc = c(0, 30),
  center_court = c(0, 47)
)

radius <- 10
B <- 100 # Number of bootstrap samples
set.seed(341)

#Calculate theta estimate at each point
calculate_theta <- function(position, data) {
  # Calculate distances and find shots within the specified radius
  dist <- sqrt((position[1] - data$LOC_X)^2 + (position[2] - data$LOC_Y)^2)
  N_0 <- which(dist <= radius)
  if (length(N_0) > 0) {
    weighted.mean(data$SHOT_MADE[N_0], w = 1 / (1 + dist[N_0]))
  }
}

#Bootstrap Sampling
bootstrap_theta <- function(position, data, B) {
  theta_bootstrap <- numeric(B)
  for (b in 1:B) {
    resampled_data <- data[sample(1:nrow(data), replace = TRUE), ]
    theta_bootstrap[b] <- calculate_theta(position, resampled_data)
  }
  theta_bootstrap
}

# Calculate theta estimates and bootstrap confidence intervals
results <- lapply(positions, function(pos) {
  theta_est <- calculate_theta(pos, nba)
  
  bootstrap_estimates <- bootstrap_theta(pos, nba, B)
  
  ci <- quantile(bootstrap_estimates, probs = c(0.025, 0.975), na.rm = TRUE)
  
  list(
    theta_estimate = theta_est,
    bootstrap_ci = ci
  )
})

# Print results
for (pos in names(results)) {
  cat("\n Position on Court:", pos, "\n")
  cat("Theta Estimate:", round(results[[pos]]$theta_estimate, 3), "\n")
  cat("95% Bootstrap CI:", round(results[[pos]]$bootstrap_ci, 3))
}

```