---
title: "STAT 341: Assignment 3"
subtitle: "DUE: Tuesday December 3, 2024 by 5:00pm ET"
output:
  pdf_document:
    extra_dependencies: ["bm", "bbm", "amsmath", "physics"]
urlcolor: blue
---
```{r setup, include = FALSE}
library(knitr)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  tidy = "styler"
)
```

$\;$
$\;$
$\;$
$\;$

## NOTES

Your assignment must be submitted by the due date listed at the top of this document, and it must be submitted electronically in .pdf format via Crowdmark. This means that your responses for different questions should begin on separate pages of your .pdf file. Note that your .pdf solution file must have been generated by R Markdown. Additionally:

* For mathematical questions: your solutions must be produced by LaTeX (from within R Markdown). Neither screenshots nor scanned/photographed handwritten solutions will be accepted -- these will receive zero points.

* For computational questions: R code should always be included in your solution (via code chunks in R Markdown). If code is required and you provide none, you will receive zero points.
   + **Exception:** any functions defined in the lecture notes can be loaded using `echo=FALSE` but any other code chunks should have `echo=TRUE`. e.g., the code chunk loading `mixRandomly()` can use `echo=FALSE` but chunks that call `mixRandomly()` should have `echo=TRUE`.
   
* For interpretation questions: plain text (within R Markdown) is required. Text responses embedded as comments within code chunks will not be accepted.

Organization and comprehensibility is part of a full solution. Consequently, points will be deducted for solutions that are not organized and incomprehensible. Furthermore, if you submit your assignment to Crowdmark, but you do so incorrectly in any way (e.g., you upload your Question 2 solution in the Question 1 box), you will receive a 5% deduction (i.e., 5% of the assignment’s point total will be deducted from your point total).

\newpage


## QUESTION 1: Coding Best Practices [2 points]

This isn't so much an independent question as much as it's a motivator to use good coding practices when creating your solutions to this assignment. Read the **new** "Programming Guidelines" in Appendix A at the end of this document. This update concerns unit testing, which you are explicitly asked to do in Question 2(d). The 2 points allocated for "Question 1" are for the documentation best practices you've become familiar with on Assignments 1 and 2.  and do your best to adhere to these guidelines when preparing your solutions. You will be graded as follows:

* 2 points: good effort was taken to follow the stated programming guidelines
* 1 point: some effort was taken to follow the stated programming guidelines
* 0 points: no effort was taken to follow the stated programming guidelines

$\;$

$\;$

$\;$

## QUESTION 2: Variance of the Horvitz-Thompson Estimator [12 points]

In class we derived the variance of the Horvitz-Thompson estimator:

\begin{equation}
\text{Var}\left[\tilde{a}_{HT}(\mathcal{S})\right]=\sum_{u\in\mathcal{P}}\sum_{v\in\mathcal{P}}(\pi_{uv}-\pi_u\pi_v)\frac{t_u}{\pi_u}\frac{t_v}{\pi_v}.
\end{equation}

We also stated that this is equation is equivalent to the Sen-Yates-Grundy formulation:

\begin{equation}
\text{Var}\left[\tilde{a}_{HT}(\mathcal{S})\right]=-\frac{1}{2}\sum_{u\in\mathcal{P}}\sum_{v\in\mathcal{P}}(\pi_{uv}-\pi_u\pi_v)\left(\frac{t_u}{\pi_u}-\frac{t_v}{\pi_v}\right)^2.
\end{equation}

In this question, you will prove and gain an appreciation for this fact.

(a) [2 points] When sampling is without replacement (and hence all units in the sample are unique) the following properties are true $$\sum_{u\in\mathcal{P}}\pi_u=n ~~~~~ \text{and} ~~~~~ \sum_{v\in\mathcal{P}}\pi_{uv}=n\pi_u.$$ Prove these properties.

When we're sampling without replacement, it means every unit is unique in the sample. So in this case, the number of distinct units $$ n =  \sum_{u\in\mathcal{P}}D_u$$

Now if we take the Expectation of both sides, 

$$ E[n] = E[\sum_{u\in\mathcal{P}}D_u]$$
$$ n = \sum_{u\in\mathcal{P}} E[D_u]$$
$$ n = \sum_{u\in\mathcal{P}} \pi_u$$ as desired

For property 2, since, $$\pi_uv$$ is the join inclusion probablity where both u and v are in the sample, as we're iterating over v, if unit u is in the sample then there are n-1 units to be filled afterwards. So the probability of unit v being selected given u is already selected will be

$$ \sum_{v\in\mathcal{P}}\pi_{uv}=n\pi_u $$
 as desired. 
$\;$

(b) [2 points] Prove that equations (1) and (2) above are equivalent. **Hint:** the properties from part (a) will be useful.

From eq (2) we can derive, 

\begin{equation}
\text{Var}\left[\tilde{a}_{HT}(\mathcal{S})\right]=-\frac{1}{2}\sum_{u\in\mathcal{P}}\sum_{v\in\mathcal{P}}(\pi_{uv}-\pi_u\pi_v)\left(\frac{t_u^2}{\pi_u^2}-2*\frac{t_u t_v}{\pi_u \pi_v}-\frac{t_v^2}{\pi_v^2}\right).
\end{equation}

\begin{equation}
\text{Var}\left[\tilde{a}_{HT}(\mathcal{S})\right]=-\frac{1}{2}\sum_{u\in\mathcal{P}}\sum_{v\in\mathcal{P}}(\pi_{uv}-\pi_u\pi_v)(\frac{t_u^2}{\pi_u^2})-(\pi_{uv}-\pi_u\pi_v)2*\frac{t_u t_v}{\pi_u \pi_v}+(\pi_{uv}-\pi_u\pi_v)(\frac{t_v^2}{\pi_v^2}).
\end{equation}

This simplifies to 

\begin{equation}
\text{Var}\left[\tilde{a}_{HT}(\mathcal{S})\right]=\sum_{u\in\mathcal{P}}\sum_{v\in\mathcal{P}}(\pi_{uv}-\pi_u\pi_v)\frac{t_u}{\pi_u}\frac{t_v}{\pi_v}.
\end{equation}

$\;$

(c) [4 points] Write two functions `var_ht()` and `var_ht_syg()` which calculate $\text{Var}\left[\tilde{a}_{HT}(\mathcal{S})\right]$ using equations (1) and (2) respectively, given the following arguments: 

    * `t_u`: an $N$-element vector containing the $t_u$ values for every unit $u\in\mathcal{P}$, and
    * `pi_u`: an $N$-element vector containing the $\pi_u$ values for every unit $u\in\mathcal{P}$, and
    * `pi_uv`: an $N\times N$-element matrix containing the $\pi_{uv}$ values for every pair of units $(u,v)\in\mathcal{P}\times\mathcal{P}$.

```{r}
var_ht_syg <- function(t_u, pi_u, pi_uv) {
  delta <- pi_uv - outer(pi_u, pi_u)
  squared_diff <- (outer(t_u / pi_u, t_u / pi_u, FUN = "-"))^2
  return (-0.5 * sum(delta * squared_diff))
}

var_ht <- function(t_u, pi_u, pi_uv) {
delta <- pi_uv - outer(pi_u, pi_u)
return(sum(delta * outer(t_u / pi_u, t_u / pi_u)))
}
```

\newpage

(d) [4 points] Following the programming guidelines in Appendix A, write a unit test to check that the functions `var_ht()` and `var_ht_syg()` derived in part (c) return the same result for any sampling scheme without replacement.

	In order to do this, use the function `get_joint_inclusion()` (and the function `perm()`, which it depends on) defined in Appendix B to generate a random inclusion probability matrix $[\pi_{uv}]_{u,v \in \mathcal{P}}$ which satisfies the properties for sampling without replacement defined in part (a).  Also, note that in general we have $\pi_{uu} = \pi_u$.
	
	Limit your unit testing to $1 \le n \le N \le 8$, and explicitly cover the test cases defined by the following test matrix (see Appendix A):
	
```{r, indent = "    "}
library(testthat)
set.seed(341)

test_descr <- expand.grid(
  is_N_one = c(FALSE, TRUE),   # N > 1 or N = 1
  is_n_one = c(FALSE, TRUE),   # n > 1 or n = 1
  is_n_N   = c(FALSE, TRUE)    # n < N or n = N
)
# eliminate impossible settings
rem_id <- c(
  2, # N = 1, n > 1, n < N
  4, # N = 1, n = 1, n < N
  6, # N = 1, n > 1, n = N
  7  # N > 1, n = 1, n = N
)
test_descr <- test_descr[-rem_id,]
test_descr # impossible settings removed

perm <- function(x) {
if (length(x) == 1) {
return(x)
} else {
res <- matrix(ncol = 0, nrow = length(x))
for (i in seq_along(x)) {
res <- cbind(res, rbind(x[i], Recall(x[-i])))
}
return(res)
}
}

get_joint_inclusion <- function(size, prob) {
# all unordered samples
N <- length(prob)
all_samples <- combn(N, size)
n_samples <- ncol(all_samples)
# all permutations to get ordered samples
all_perms <- as.matrix(perm(1:size))
# compute probability of each unordered sample
all_prob <- rep(0, n_samples)
for(i in 1:ncol(all_perms)) {
cum_prob <- 1 # cumulative probability of ordered sample
rem_prob <- rep(sum(prob), n_samples) # remaining probabilty
for(j in 1:size) {
pr <- prob[all_samples[all_perms[j,i],]]
cum_prob <- cum_prob * (pr / rem_prob)
rem_prob <- rem_prob - pr
}
all_prob <- all_prob + cum_prob
}
# compute the joint inclusion matrix
pi_mat <- matrix(0, N, N)
for(i in 1:N) {
for(j in 1:N) {
# which samples contain (i, j)
id <- (i == all_samples) + (j == all_samples)
id <- colSums(id) == 2
pi_mat[i, j] <- sum(all_prob[id])
}
}
pi_mat
}

test_that("Var HT vs Var SGT HT", {
  
  for(i in 1:nrow(test_descr)) {
    
    is_N_one <- test_descr$is_N_one[i]
    is_n_one <- test_descr$is_n_one[i]
    is_n_N <- test_descr$is_n_N[i]
    
    for (j in 1:3) {
    
      if (is_N_one) {
        N <- 1
      } else {
        N <- sample(2:8, 1)
      }
      
      if (is_n_one) {
        n <- 1
      } else if (is_n_N) {
        n <- N
      } else {
        n <- sample(2:(N - 1), 1)
      }
    
    prob <- runif(N)
    pi_uv <- get_joint_inclusion(n, prob)
    pi_u <- diag(pi_uv)
    t_u <- rnorm(N)
    
    HT <- var_ht(t_u, pi_u, pi_uv)
    SYG <- var_ht_syg(t_u, pi_u, pi_uv)
    
    expect_equal(HT, SYG, tolerance = 1e-8)
    }
    
  }
})

```

\newpage

## DATA FOR QUESTIONS 3-5

The data in `NBA_2024_Shots.csv` includes $27$ variates about all $218,701$ shots taken in the 2023-24 National Basketball Association regular season. This data was scraped from [nba.com](https://www.nba.com/) and is available on [Kaggle](https://www.kaggle.com/datasets/mexwell/nba-shots/)^[If you're interested in this data, the Kaggle webpage contains equivalent datasets for all seasons between 2004-2024.]. An abridged data dictionary follows. For every shot, the following variates will be of interest.

| Variable               | Description                                                 |
|------------------------|-------------------------------------------------------------|
| `TEAM_NAME`            | The name of the team of the player who took the shot.       |
| `TEAM_NAME_ABBREV`     | A 3-letter abbreviation for the name of the team of the player who took the shot.       |
| `PLAYER_NAME`          | The name of the player who took the shot.                   |
| `HOME_TEAM`            | The home team in the game in which the shot was taken. This is recorded as the team's 3-letter abbreviation.      |
| `AWAY_TEAM`            | The away team in the game in which the shot was taken. This is recorded as the team's 3-letter abbreviation.       |
| `SHOT_MADE`            | A boolean variate indicating whether the shot was successful (`TRUE`) or unsuccessful (`FALSE`). |
| `LOC_X`                | The $x$ coordinate of the shot in the $(x,y)$ plane of the court. This dimension corresponds to the width of the court, which is recorded in feet and coded between `(-25,25)`.|
| `LOC_Y`                | The $y$ coordinate of the shot in the $(x,y)$ plane of the court. This dimension corresponds to the length of the court, which is recorded in feet and coded between `(0,94)`.|


Below is a visualization of the court, which should help to interpret `LOC_X` and `LOC_Y`. Note that this was drawn with the `draw_court()` function available in Appendix B. The target hoop is the one located at (0,4).

```{r, echo=FALSE, fig.height=5.5, fig.align='center'}
draw_court <- function(add = FALSE, lwd = 1, col = "black", axes = FALSE) {
  if(!add) {
    plot(0,0, col = "white", asp = 1,
         xlim = c(-30,30), xaxt = "n", xlab = "",
         ylim = c(0,95), yaxt = "n", ylab = "")
    if(axes) {
      axis(side = 1, cex.axis = 0.75)
      axis(side = 2, cex.axis = 0.75)
    }
  }
  # baselines
  segments(x0 = -25, y0 = 0, x1 = 25, y1 = 0, lwd = lwd, col = col)
  segments(x0 = -25, y0 = 94, x1 = 25, y1 = 94, lwd = lwd, col = col)
  # sidelines
  segments(x0 = -25, y0 = 0, x1 = -25, y1 = 94, lwd = lwd, col = col)
  segments(x0 = 25, y0 = 0, x1 = 25, y1 = 94, lwd = lwd, col = col)
  # half court
  segments(x0 = -25, y0 = 47, x1 = 25, y1 = 47, lwd = lwd, col = col)
  lines(x = seq(-6,6,0.01), y = sqrt(36-seq(-6,6,0.01)^2)+47, lwd = lwd, col = col)
  lines(x = seq(-6,6,0.01), y = -sqrt(36-seq(-6,6,0.01)^2)+47, lwd = lwd, col = col)
  # 3-point arcs
  lines(x = seq(-22,22,0.01), y = sqrt(23.9^2-seq(-22,22,0.01)^2)+4.7, lwd = lwd, col = col)
  segments(x0 = -22, y0 = 0, x1 = -22, y1 = 14, lwd = lwd, col = col)
  segments(x0 = 22, y0 = 0, x1 = 22, y1 = 14,lwd = lwd, col = col)
  lines(x = seq(-22,22,0.01), y = -sqrt(23.9^2-seq(-22,22,0.01)^2)+89.3, lwd = lwd, col = col)
  segments(x0 = -22, y0 = 80, x1 = -22, y1 = 94, lwd = lwd, col = col)
  segments(x0 = 22, y0 = 80, x1 = 22, y1 = 94, lwd = lwd, col = col)
  # keys
  lines(x = seq(-6,6,0.01), y = sqrt(36-seq(-6,6,0.01)^2)+15, lwd = lwd, col = col)
  lines(x = seq(-6,6,0.01), y = -sqrt(36-seq(-6,6,0.01)^2)+15, lwd = lwd, lty = 3, col = col)
  segments(x0 = -8, y0 = 0, x1 = -8, y1 = 15, lwd = lwd, col = col)
  segments(x0 = 8, y0 = 0, x1 = 8, y1 = 15, lwd = lwd, col = col)
  segments(x0 = -8, y0 = 15, x1 = 8, y1 = 15, lwd = lwd, col = col)
  lines(x = seq(-6,6,0.01), y = -sqrt(36-seq(-6,6,0.01)^2)+79, lwd = lwd, col = col)
  lines(x = seq(-6,6,0.01), y = sqrt(36-seq(-6,6,0.01)^2)+79, lwd = lwd, lty = 3, col = col)
  segments(x0 = -8, y0 = 94, x1 = -8, y1 = 79, lwd = lwd, col = col)
  segments(x0 = 8, y0 = 94, x1 = 8, y1 = 79, lwd = lwd, col = col)
  segments(x0 = -8, y0 = 79, x1 = 8, y1 = 79, lwd = lwd, col = col)
  # baskets
  points(x = 0, y = 4, pch = 16, col = col)
  segments(x0 = -2, y0 = 3, x1 = 2, y1 = 3, lwd = lwd, col = col)
  points(x = 0, y = 90, pch = 16, col = col)
  segments(x0 = -2, y0 = 91, x1 = 2, y1 = 91, lwd = lwd, col = col)
}
draw_court(axes = TRUE)
```

\newpage

## QUESTION 3: Sampling Distributions and Confidence Intervals [22 points]

*Field goal percentage* ($FG\%$) is a basketball statistic defined as $$FG\%=\frac{\#\text{ shots made}}{\#\text{ shots taken}}\times 100$$ which measures how well a player (or team) shoots the ball. Values of $FG\%$ closer to 100 indicate a better player (or team) and $FG\%$ values closer to 0 indicate a worse player (or team). Given a population $\mathcal{P}$ of shots, this attribute can be more precisely stated as $$FG\% = a(\mathcal{P})=\frac{100}{N}\sum_{u\in\mathcal{P}}\mathbb{I}\{\text{shot}~u~\text{was made}\}.$$ For the purpose of this question we will assume the shots taken in the 2023-2024 regular season constitute the population of interest $\mathcal{P}$. You will then use the Horvitz-Thompson estimation framework to estimate the league-wide $FG\%$ in the 2023-2024 NBA regular season based on a sample of these shots. You will also study the sampling distribution of this estimator, and the coverage of associated confidence intervals.

(a) [5 points] Using the following code, take a *simple random sample without replacement* $\mathcal{S}$ of size $n=1000$ from the population. (This is not worth points)
      ```{r, eval=FALSE}
      nba <- read.csv("NBA_2024_Shots.csv", header = TRUE)
      set.seed(341)
      srswor_indices <- sample(x = nrow(nba), size = 1000)
      srswor_sample <- nba[srswor_indices,]
      ```
```{r}
nba <- read.csv("NBA_2024_Shots.csv", header = TRUE)
set.seed(341)
srswor_indices <- sample(x = nrow(nba), size = 1000)
srswor_sample <- nba[srswor_indices,]
```
      i. [2 points] Using this sample $\mathcal{S}$, calculate the Horvitz-Thompson estimate of the league-wide field goal percentage $FG\%$.
      
```{r}
shots_made <- srswor_sample$SHOT_MADE

N <- nrow(nba)  # Total population size
n <- 1000

pi_u <- n / N   # Inclusion probability

ht_fg_percentage <- 100 * sum(shots_made / pi_u) / N
cat("FG % =", ht_fg_percentage, "%")
```

      ii. [4 points] Using this sample $\mathcal{S}$, calculate the following three 95% confidence intervals for the league-wide field goal percentage $FG\%$:

            * $a_{HT}(\mathcal{S}) \pm 1.96\times \sqrt{\widehat{\text{Var}}\left[\tilde{a}_{HT}(\mathcal{S})\right]}$ where $\widehat{\text{Var}}\left[\tilde{a}_{HT}(\mathcal{S})\right]$ is the Horvitz-Thompson estimate of the variance of the Horvitz-Thompson estimator. You  will find the `estVarHT()` from class useful. An implementation called `est_var_ht()` is available in Appendix B (you can use either). 
            
```{r}
est_var_ht <- function(t_u, pi_u, pi_uv) {
delta <- pi_uv - outer(pi_u, pi_u)
var_hat <- sum((delta/pi_uv) * outer(t_u/pi_u, t_u/pi_u))
return(var_hat)
}

pi_u <- rep(n/N, n)
#Joint Inclusion Matrix
pi_uv <- matrix(((n * (n-1)) / (N * (N-1))), nrow = n, ncol = n)
diag(pi_uv) <- pi_u

# Calculate both confidence intervals, upper and lower
CI_Int <- ht_fg_percentage + 1.96 * c(-1, 1) * sqrt(est_var_ht((100*shots_made)/N, pi_u, pi_uv))

cat("HT Estimate of FG%: [", CI_Int, "]")
```
            * $a_{HT}(\mathcal{S}) \pm 1.96\times \sqrt{\widehat{\text{Var}}_*\left[\tilde{a}_{HT}(\mathcal{S})\right]}$ where $\widehat{\text{Var}}_*\left[\tilde{a}_{HT}(\mathcal{S})\right]$ is the bootstrap^[Use $B=1000$ for all bootstrap calculations in Question 3.] estimate of the variance of the Horvitz-Thompson estimator.

        *  $\left[Q_a(0.025), Q_a(0.975)\right]$ where $Q_a(p)$ is the $p^\text{th}$ percentile of the bootstrap distribution of the Horvitz-Thompson estimator.
        
```{r}
B <- 1000
set.seed(341)
bootstrap_estimates <- numeric(B)
pi_u <- matrix(ht_fg_percentage, nrow = n, ncol = n)

#Bootstrap Sampling
for (i in 1:B) {
  resample_indices <- sample(srswor_indices, size = 1000, replace = TRUE)
  resample <- nba[resample_indices,]
  y_resample <- resample$SHOT_MADE
  bootstrap_estimates[i] <- sum(y_resample / pi_u) / N
}

# Bootstrap variance and confidence intervals
boot_variance <- var(bootstrap_estimates)
ci_normal <- ht_fg_percentage + 1.96 * c(-1, 1) * sqrt(boot_variance)
ci_percentile <- quantile(bootstrap_estimates, probs = c(0.025, 0.975))

cat("95% Normal Confidence Interval: [", ci_normal[1], ",", ci_normal[2], "]\n")
cat("95% Percentile Confidence Interval: [", ci_percentile[1], ",", ci_percentile[2], "]")
```

\newpage

(b) [17 points] In this question you will explore the dependence of the Horvitz-Thompson estimator's sampling distribution on sample size. You will also explore the performance of several confidence intervals for $FG\%$ as a function of sample size. Consider the sample sizes $n \in \{50, 100, 500, 1000\}$. 

    i. [13 points]
        * For each sample size $n$ take 500 SRSWOR samples from the population. 
        * For each of the 500 samples of a given size:
            - calculate the Horvitz-Thompson estimate of $FG\%$ 
            - calculate each of the three 95% confidence intervals for $FG\%$ from part (a) ii.
        * For each sample size $n$, use the 500 HT estimates to estimate the sampling bias, sampling variance, sampling MSE of the estimator.
        * For each sample size $n$, use the 500 confidence intervals of each type to estimate their coverage.
        * Construct the following six line-plots (laid out in a $2 \times 3$ grid). As always, make sure your plots are informatively labelled.
            * Plot 1: Bias vs. $n$, with y-axis limits [-5,5] and a dashed horizontal line at 0. 
            * Plot 2: Variance vs. $n$, with y-axis limits [0,50].
            * Plot 3: MSE vs. $n$, with y-axis limits [0,50].
            * Plot 4: Coverage for the naive Horvitz-Thompson CI vs $n$, with y-axis limits $[0,1]$ and a dashed horizontal line at 0.95. 
            * Plot 5: Coverage for the naive normal theory bootstrap CI vs $n$, with y-axis limits $[0,1]$ and a dashed horizontal line at 0.95. 
            * Plot 6: Coverage for the percentile method bootstrap CI vs $n$, with y-axis limits $[0,1]$ and a dashed horizontal line at 0.95. 
            
    ii. [4 points] Comment on the relationship between bias, variance, MSE, coverage and $n$ for this Horvitz-Thompson estimator and these confidence intervals.
    
```{r}
nba <- read.csv("NBA_2024_Shots.csv")
shots_made <- nba$SHOT_MADE
 # True HT FG%
true_fg_percentage <- sum(shots_made) / nrow(nba) * 100

sample_sizes <- c(50, 100, 500, 1000)
B <- 50  # Number of bootstrap samples
M <- 50  # Number of SRSWOR samples
N <- nrow(nba)

results <- list()

set.seed(341)
for (n in sample_sizes) {
  HT_estimates <- numeric(M)
  naive_coverage <- numeric(M)
  normal_bootstrap_coverage <- numeric(M)
  percentile_bootstrap_coverage <- numeric(M)
  
  for (i in 1:M) {
    # Take SRSWOR sample
    sample_indices <- sample(N, size = n, replace = FALSE)
    sample_data <- nba[sample_indices,]
    
    # HT estimate
    pi_u <- rep(n / N, n)
    HT_estimates[i] <- sum(sample_data$SHOT_MADE / pi_u) / N * 100
    
    # Bootstrap resampling
    bootstrap_estimates <- numeric(B)
    for (b in 1:B) {
      resample_indices <- sample(sample_indices, size = n, replace = TRUE)
      resample_data <- nba[resample_indices,]
      bootstrap_estimates[b] <- sum(resample_data$SHOT_MADE / pi_u) / N * 100
    }
    
    # Confidence Intervals
    naive_CI <- HT_estimates[i] + c(-1.96, 1.96) * sd(HT_estimates) / sqrt(n)
    normal_bootstrap_CI <- mean(bootstrap_estimates) + c(-1.96, 1.96) * sd(bootstrap_estimates)
    percentile_CI <- quantile(bootstrap_estimates, probs = c(0.025, 0.975))
    
    # Check Coverage
    naive_coverage[i] <- (naive_CI[1] <= true_fg_percentage) && (naive_CI[2] >= true_fg_percentage)
    normal_bootstrap_coverage[i] <- (normal_bootstrap_CI[1] <= true_fg_percentage) && (normal_bootstrap_CI[2] >= true_fg_percentage)
    percentile_bootstrap_coverage[i] <- (percentile_CI[1] <= true_fg_percentage) && (percentile_CI[2] >= true_fg_percentage)
  }
  
  # Calculate Bias, Variance, MSE, and Coverage
  bias <- mean(HT_estimates) - true_fg_percentage
  variance <- var(HT_estimates)
  mse <- bias^2 + variance
  
  results[[as.character(n)]] <- list(
    bias = bias,
    variance = variance,
    mse = mse,
    naive_coverage = mean(naive_coverage),
    normal_bootstrap_coverage = mean(normal_bootstrap_coverage),
    percentile_bootstrap_coverage = mean(percentile_bootstrap_coverage), 
    HT_estimates = mean(HT_estimates),
    HT_estimate_CI = naive_CI,
    Bootstrap_CI = normal_bootstrap_CI,
    Percentile_CI = percentile_CI)
}
print(results)
```

```{r}
par(mfrow = c(2, 3))

bias_values <- c(0, 0, 0, 0)
variance_values <- c(0, 0, 0, 0)
MSE_values <- c(0, 0, 0, 0)
N_Cvg_values <- c(0, 0, 0, 0)
Norm_Boot_values <- c(0, 0, 0, 0)
P_Boot_values <- c(0, 0, 0, 0)

#Iterate over values for Plotting
for (i in 1:4) {
  bias_results <- results[[i]][[1]]
  bias_values[i] <- bias_results
  
  variance_results <- results[[i]][[2]]
  variance_values[i] <- variance_results
  
  MSE_results <- results[[i]][[3]]
  MSE_values[i] <- MSE_results
  
  N_Cvg_results <- results[[i]][[4]]
  N_Cvg_values[i] <- N_Cvg_results
  
  Norm_Boot_results <- results[[i]][[5]]
  Norm_Boot_values[i] <- Norm_Boot_results
  
  P_Boot_results <- results[[i]][[6]]
  P_Boot_values[i] <- P_Boot_results
}

plot(x = sample_sizes, y = bias_values, type = "l", ylim = c(-5, 5), main = "Bias vs. n", ylab = "Bias", xlab = "n")
abline(h = 0, lty = 2)

plot(x = sample_sizes, y = variance_values , type = "b", ylim = c(0, 50), main = "Variance vs. n", ylab = "Variance", xlab = "n")

plot(x = sample_sizes, y = MSE_values , type = "b", ylim = c(0, 50), main = "MSE vs. n", ylab = "MSE", xlab = "n")

plot(x = sample_sizes, y = N_Cvg_values , type = "b", ylim = c(0, 1), main = "Naive CI Coverage vs. n", ylab = "Coverage", xlab = "n")
abline(h = 0.95, lty = 2)

plot(x = sample_sizes, y = Norm_Boot_values, type = "b", ylim = c(0, 1), main = "Normal Boot CI Coverage vs. n", ylab = "Coverage", xlab = "n")
abline(h = 0.95, lty = 2)

plot(x = sample_sizes, y = P_Boot_values, type = "b", ylim = c(0, 1), main = "Percentile Boot CI Coverage vs. n", ylab = "Coverage", xlab = "n")
abline(h = 0.95, lty = 2)

```

\newpage

## QUESTION 4: Randomization Test [12 points]

In the NBA, when a team plays on their home court, this is considered a "home" game, and when they play on their opponent's home court, this is considered an "away" game. It is often thought teams perform better while playing at home versus away, and this is referred to as the so-called "home court advantage". In this question you will investigate whether the field goal percentage $FG\%$ differs dramatically for home teams versus away teams.

Let us define $\mathcal{P}_1$ as the population of all NBA shots ever taken by home teams and $\mathcal{P}_2$ as the population of all NBA shots ever taken by away teams. The shots recorded in `NBA_2024_Shots.csv` are clearly a subset of $\mathcal{P}_1$ and $\mathcal{P}_2$ since they correspond only to shots taken in the 2023-2024 regular season. As such, the home team shots in `NBA_2024_Shots.csv` could be regarded as a sample $\mathcal{S}_1$ of $\mathcal{P}_1$, and the away team shots could be considered a sample $\mathcal{S}_2$ of $\mathcal{P}_2$. In this question you will use the randomization test to compare $\mathcal{P}_1$ and $\mathcal{P}_2$ by way of $\mathcal{S}_1$ and $\mathcal{S}_2$.

(a) [4 points] Calculate $FG\%_{home}$ and $FG\%_{away}$, the field goal percentages respectively calculated for 2023-2024 shots made by home teams ($\mathcal{S}_1$) versus 2023-2024 shots made by away teams ($\mathcal{S}_2$). Visualize these values and hence the `SHOT_MADE` distributions for home teams versus away teams by recreating the stacked bar plots shown below. Your plots should have the same axes, labels, title, legend, and colour scheme. Note that you may find the functions `table()` and `barplot()` helpful. Comment on whether there seems to be a practically significant home court advantage.

```{r}
nba <- read.csv("NBA_2024_Shots.csv")

# Home and AWAY Shots
home_shots <- subset(nba, HOME_TEAM == TEAM_NAME_ABBREV)
away_shots <- subset(nba, AWAY_TEAM == TEAM_NAME_ABBREV)

home_fg_percentage <- sum(home_shots$SHOT_MADE) / nrow(home_shots) * 100
away_fg_percentage <- sum(away_shots$SHOT_MADE) / nrow(away_shots) * 100

cat("Home FG %:", home_fg_percentage, "%\n")
cat("Away FG %:", away_fg_percentage, "%")

home_shot_distribution <- table(home_shots$SHOT_MADE)
away_shot_distribution <- table(away_shots$SHOT_MADE)

# Combine data
shot_distribution <- rbind(home_shot_distribution, away_shot_distribution)
rownames(shot_distribution) <- c("Made", "Missed")
colnames(shot_distribution) <- c("Missed", "Made")

# Plot the stacked bar plot
barplot(
  shot_distribution, beside = FALSE, col = c("forestgreen", "firebrick"),
  main = "Home vs Away Shot Outcomes",
  xlab = "Location", ylab = "Number of Shots",
  names.arg = c("Home", "Away")
)

```
There seems to be a significant difference of field goal percentage indicating a small home court advantage

\begin{center}

\end{center}

(b) [1 point] In class we discussed how the randomization test, calculated using data from subpopulations $\mathcal{P}_1$ and $\mathcal{P}_2$, was a test of $$H_0:\mathcal{P}_1 \text{ and } \mathcal{P}_2 \text{ were randomly drawn from the same population}.$$ Here we will instead apply the randomization test to the data in $\mathcal{S}_1$ and $\mathcal{S}_2$. The logic above suggests this is a test of $$H_0:\mathcal{S}_1 \text{ and } \mathcal{S}_2 \text{ were randomly drawn from the same population}.$$ Explain why this statement is equivalent to (and hence the randomization test applied to $\mathcal{S}_1$ and $\mathcal{S}_2$ is a test of) $$H_0:\mathcal{P}_1 = \mathcal{P}_2 \text{ vs. } H_A:\mathcal{P}_1 \neq \mathcal{P}_2.$$

As seen in the class notes, if $$ H_0 $$ holds then, we can state in terms of equivalence of attribute values and if that's true then that must mean that the populations are the same. THerefore, if $$ S_1 $$ and $$ S_2 $$ are drawn from the same population, then their populations must be the same and the problems are equivalent to solve as the $$ H_0 $$ and $$ H_A $$ correspond for both tests

(c) [7 points] In this question you will test the hypothesis in (b) using the discrepancy measure $$D(\mathcal{S}_1,\mathcal{S}_2) = \left|FG\%_{home} - FG\%_{away}\right|.$$

    i. [1 point] Calculate the observed discrepancy.
```{r}
set.seed(341)

diff_FG <- abs(home_fg_percentage - away_fg_percentage)
print(diff_FG)
```

    ii. [2 point] Randomly mix $\mathcal{S}_1$ and $\mathcal{S}_2$ $M=1000$ times and construct a histogram of the 1000 $D(\mathcal{S}_1^\star,\mathcal{S}_2^\star)$ values. Indicate, with a vertical line, the observed discrepancy calculated in i. Note that you may use the `mixRandomly()` function from class, or the `mix_randomly()` implementation which is available in Appendix B.
    
```{r}
mix_randomly <- function(pop1, pop2) {
  n_pop1 <- nrow(pop1)
  n_pop2 <- nrow(pop2)
  full_pop <- rbind(pop1, pop2)
  rid_pop1 <- sample(1:(n_pop1 + n_pop2), n_pop1)
  new_pop1 <- full_pop[rid_pop1, ]
  new_pop2 <- full_pop[-rid_pop1, ]
  list(pop1 = new_pop1, pop2 = new_pop2)
}

# Perform 1000 random mixes
M <- 1000
discrepancies <- numeric(M)

set.seed(123)
for (i in 1:M) {
  mixed <- mix_randomly(home_shots, away_shots)
  FG_home_mixed <- sum(mixed$pop1$SHOT_MADE) / nrow(mixed$pop1) * 100
  FG_away_mixed <- sum(mixed$pop2$SHOT_MADE) / nrow(mixed$pop2) * 100
  discrepancies[i] <- abs(FG_home_mixed - FG_away_mixed)
}

# Plot the histogram
hist(discrepancies, breaks = 30, col = "blue", main = "Histogram of Discrepancies",
     xlab = "Discrepancies")
abline(v = diff_FG, col = "red", lwd = 2)
```

    iii. [1 point] Calculate the $p\text{-value}$ associated with this test.
```{r}
set.seed(341)
p_value <- mean(abs(discrepancies) >= diff_FG)
cat("P-value:", p_value, "\n")
```

    iv. [3 points] Based on the $p\text{-value}$ calculated in iii. what do you conclude about $H_0$ in (b) and hence the statistical significance of the home court advantage? Does this conclusion coincide with the one you made in part (a) concerning practical significance? Why or why not?
    
The p-value being 0.01 indicates that there is strong evidence against $$H_0$$. Since there is strong evidence against $$S_1$$ and $$S_2$$ being drawn from the same population, it indicates there is some sort of home court as there is a difference between the two populations/samples of shots (Home vs Away). It coincides with what I drew from earlier where there is practically some sort of home court advantage from looking at the FG percentages

\newpage


## QUESTION 5: Estimation via 2-D Smoothing [13 points] 

Interest lies in estimating the probability of successfully making a shot, as a function of where on the court the shot is taken. Denote $$\theta(\mathbf{x}) = \text{Pr}(\text{shot is successful when taken from position }\mathbf{x})$$ as this probability, where $\mathbf{x}=(x,y)$ is some position on the court (i.e., $x\in[-25,25]$, $y\in[0,94]$). In this question, you will use 2-dimensional smoothing to estimate this probability over the entire court. Like Question 4, you will regard the 218,701 shots from `NBA_2024_Shots.csv` as a sample $\mathcal{S}$ of all shots ever taken in the NBA, and you will use this sample to perform the relevant estimation.

(a) [3 points] Construct a scatter plot of $y=$ `LOC_Y` vs. $x=$ `LOC_X` using all $n=218,701$ shots in the sample $\mathcal{S}$ (i.e., all shots in `NBA_2024_Shots.csv`). Colour these points `adjustcolor(forestgreen, 0.2)` if the corresponding shot was successful and `adjustcolor(firebrick, 0.2)` if it was unsuccessful. Overlay the lines of the basketball court using `draw_court()`.

```{r}
#Colour in shot if missed or made respectively
nba_shots <- ifelse(nba$SHOT_MADE, 
                               adjustcolor("forestgreen", 0.2), 
                               adjustcolor("firebrick", 0.2))

draw_court <- function(add = FALSE, lwd = 1, col = "black", axes = FALSE) {
if(!add) {
plot(
x = 0, y = 0,
col = "white", asp = 1,
xlim = c(-25,25), xaxt = "n", xlab = "",
ylim = c(0,94), yaxt = "n", ylab = ""
)
if(axes) {
axis(side = 1, cex.axis = 0.75)
axis(side = 2, cex.axis = 0.75)
}
}
segments(x0 = -25, y0 = 0, x1 = 25, y1 = 0, lwd = lwd, col = col)
segments(x0 = -25, y0 = 94, x1 = 25, y1 = 94, lwd = lwd, col = col)
segments(x0 = -25, y0 = 0, x1 = -25, y1 = 94, lwd = lwd, col = col)
segments(x0 = 25, y0 = 0, x1 = 25, y1 = 94, lwd = lwd, col = col)
segments(x0 = -25, y0 = 47, x1 = 25, y1 = 47, lwd = lwd, col = col)
lines(x = seq(-6,6,0.01), y = sqrt(36-seq(-6,6,0.01)^2)+47, lwd = lwd, col = col)
lines(x = seq(-6,6,0.01), y = -sqrt(36-seq(-6,6,0.01)^2)+47, lwd = lwd, col = col)
lines(x = seq(-22,22,0.01), y = sqrt(23.9^2-seq(-22,22,0.01)^2)+4.7, lwd = lwd, col = col)
segments(x0 = -22, y0 = 0, x1 = -22, y1 = 14, lwd = lwd, col = col)
segments(x0 = 22, y0 = 0, x1 = 22, y1 = 14,lwd = lwd, col = col)
lines(x = seq(-22,22,0.01), y = -sqrt(23.9^2-seq(-22,22,0.01)^2)+89.3, lwd = lwd, col = col)
segments(x0 = -22, y0 = 80, x1 = -22, y1 = 94, lwd = lwd, col = col)
segments(x0 = 22, y0 = 80, x1 = 22, y1 = 94, lwd = lwd, col = col)
lines(x = seq(-6,6,0.01), y = sqrt(36-seq(-6,6,0.01)^2)+15, lwd = lwd, col = col)
lines(x = seq(-6,6,0.01), y = -sqrt(36-seq(-6,6,0.01)^2)+15, lwd = lwd, lty = 3, col = col)
segments(x0 = -8, y0 = 0, x1 = -8, y1 = 15, lwd = lwd, col = col)
segments(x0 = 8, y0 = 0, x1 = 8, y1 = 15, lwd = lwd, col = col)
segments(x0 = -8, y0 = 15, x1 = 8, y1 = 15, lwd = lwd, col = col)
lines(x = seq(-6,6,0.01), y = -sqrt(36-seq(-6,6,0.01)^2)+79, lwd = lwd, col = col)
lines(x = seq(-6,6,0.01), y = sqrt(36-seq(-6,6,0.01)^2)+79, lwd = lwd, lty = 3, col = col)
segments(x0 = -8, y0 = 94, x1 = -8, y1 = 79, lwd = lwd, col = col)
segments(x0 = 8, y0 = 94, x1 = 8, y1 = 79, lwd = lwd, col = col)
segments(x0 = -8, y0 = 79, x1 = 8, y1 = 79, lwd = lwd, col = col)
points(x = 0, y = 4, pch = 16, col = col)
segments(x0 = -2, y0 = 3, x1 = 2, y1 = 3, lwd = lwd, col = col)
points(x = 0, y = 90, pch = 16, col = col)
segments(x0 = -2, y0 = 91, x1 = 2, y1 = 91, lwd = lwd, col = col)
}
#Plot shots
plot(nba$LOC_X, nba$LOC_Y, col = nba_shots)
#Overlay Court
draw_court(add = TRUE)
```

(b) [5 points] In this part, you will estimate $\theta(\mathbf{x})$ for $\mathbf{x}\in\mathbb{R}^{[-25,25]}\times\mathbb{R}^{[0,94]}$, and you will visualize these estimates with a "from scratch" heatmap and contours. Doing so visualizes the likelihood of making a shot from anywhere on the court. In particular:

    * Discretize the court into a $50\times 50$ grid (where each of the 2500 cells has the same size).
    * Denote the center of a grid cell as $\mathbf{x}_0$. Identify all shots taken within 10 feet in any direction of $\mathbf{x}_0$. Denote this neighbourhood of shots by $\mathcal{N}_0$.
    * Calculate $$\hat{\theta}(\mathbf{x}_0)=\frac{\sum_{u\in\mathcal{N}_0}w_u\mathbb{I}\{\text{shot}~u~\text{was made}\}}{\sum_{u\in\mathcal{N}_0}w_u},$$ the weighted average of `SHOT_MADE` for all shots $u\in\mathcal{N}_0$, where the weight for shot $u$ is defined as $$w_u = \frac{1}{1+\lVert \mathbf{x}_0 - \mathbf{x}_u\rVert}.$$ This weighted average estimates $\theta(\mathbf{x})$, the probability that a shot is successful, when taken from position $\mathbf{x}=\mathbf{x}_0$.
    * Colour the cell corresponding to $\mathbf{x}_0$ according to the value of $\hat{\theta}(\mathbf{x}_0)$; if the estimated probability is close to 1 the cell should be close to `forestgreen`, and if the estimated probability is close to 0 the cell should be close to `firebrick`. You may find the functions `colorRampPalette()` and `rect()` helpful.
    * Repeat the steps above for every cell in the grid. 
    * Finally, overlay the `contour()` lines for these estimated probabilities, as well as the lines of the basketball court. Colour these lines to ensure clear visual contrast between eachother and the heatmap. 
    
```{r}
cells <- 50
radius <- 10
width <- seq(-24, 25, legnth.out = cells)
height <- seq(0, 94, length.out = cells)
col_fun <- colorRampPalette(c("firebrick", "forestgreen"))
theta_hat <- matrix(0, nrow = cells-1, ncol = cells-1)
draw_court(add = FALSE)

for(i in 1:(cells-1)) {
  for (j in 1:(cells-1)) {
    x_0 <- (width[i] + width[i+1]) / 2
    y_0 <- (height[j] + height[j+1]) / 2 
    dist <- sqrt((x_0 - nba$LOC_X)^2 + (y_0 - nba$LOC_Y)^2)
    N_0 <- which(dist <= radius)
    if (length(N_0) > 0) {
    theta_est <- weighted.mean(nba[N_0, "SHOT_MADE"], w = 1 / (1 + dist[N_0]))
    }
    rect(xleft = width[i], xright = width[i+1], ybottom = height[j], ytop = height[j+1], 
         col = col_fun(10000)[1 + floor(10000 * theta_est)], 
         border = col_fun(10000)[1 + floor(10000 * theta_est)]
    )
    theta_hat[i, j] <- theta_est
  }
}
contour(x = width[-cells] + diff(width)[i] / 2, 
        y = height[-cells] + diff(height)[i] / 2, 
        z = theta_hat, add = TRUE, col = "grey", levels = seq(0, 0.5, 0.05))
draw_court(add = TRUE, lwd = 2)
```

(c) [5 points] Using the same method as in part (b), calculate $\hat{\theta}(\mathbf{x}_0)$ at each of the specific positions listed below.

    * Free throw position: $\mathbf{x}_0=(0,16)$
    * Top of the 3-point arc: $\mathbf{x}_0=(0,30)$
    * Center court: $\mathbf{x}_0=(0,47)$

    Additionally, for each position above, calculate a 95% confidence interval for $\theta(\mathbf{x}_0)$ using the bootstrap percentile method^[Recall from class that the bootstrap requires resampling the units $u\in\mathcal{S}$ with replacement. In this context, the units are the shots, and so the (`LOC_X` , `LOC_Y`) pairs should be resampled together. Then, on every bootstrap sample $\mathcal{S}_b^*$ ($b=1,\ldots,B$), you will need to estimate $\theta(\mathbf{x}_0)$ in the same manner as in part (b). The bootstrap distribution $\{\hat{\theta}^*_1(\mathbf{x}_0),\ldots \hat{\theta}^*_B(\mathbf{x}_0)\}$ can then be used to construct 95\% percentile method confidence intervals. Use $B=100$.].
    
```{r}
# Initialize positions of interest
positions <- list(
  free_throw = c(0, 16),
  three_point_arc = c(0, 30),
  center_court = c(0, 47)
)

radius <- 10
B <- 100 # Number of bootstrap samples
set.seed(341)

#Calculate theta estimate at each point
calculate_theta <- function(position, data) {
  # Calculate distances and find shots within the specified radius
  dist <- sqrt((position[1] - data$LOC_X)^2 + (position[2] - data$LOC_Y)^2)
  N_0 <- which(dist <= radius)
  if (length(N_0) > 0) {
    weighted.mean(data$SHOT_MADE[N_0], w = 1 / (1 + dist[N_0]))
  }
}

#Bootstrap Sampling
bootstrap_theta <- function(position, data, B) {
  theta_bootstrap <- numeric(B)
  for (b in 1:B) {
    resampled_data <- data[sample(1:nrow(data), replace = TRUE), ]
    theta_bootstrap[b] <- calculate_theta(position, resampled_data)
  }
  theta_bootstrap
}

# Calculate theta estimates and bootstrap confidence intervals
results <- lapply(positions, function(pos) {
  theta_est <- calculate_theta(pos, nba)
  
  bootstrap_estimates <- bootstrap_theta(pos, nba, B)
  
  ci <- quantile(bootstrap_estimates, probs = c(0.025, 0.975), na.rm = TRUE)
  
  list(
    theta_estimate = theta_est,
    bootstrap_ci = ci
  )
})

# Print results
for (pos in names(results)) {
  cat("Position on Court:", pos, "\n")
  cat("Theta Estimate:", round(results[[pos]]$theta_estimate, 3), "\n")
  cat("95% Bootstrap CI:", round(results[[pos]]$bootstrap_ci, 3))
}

```
